Bullet Data Preparation for LIME Diagnostics Paper

### 2020-09-28 code/02-data-preparation.Rmd
finished updating preparation code based on CCFs_withlands training data

1. Rename the label variables using numbers instead of letters (e.g. `study.x` and `study.y` become `study1` and `study2`, respectively). 
2. Select only the variables of interest for the study: signature ID labels (`study1`, `barrel1`, `bullet1`, `land1`, `study2`, `barrel2`, `bullet2`, and `land2`), features used to train the `rtrees` random forest model (`ccf`, `rough_cor`, `D`, `sd_D`, `matches`, `mismatches`, `cms`, `non_cms`, and `sum_peaks`), and the variable identifying whether the two signatures are fired from the same gun (`same_source`).
3. Convert all categorical variables to characters.
4. Change the study labels to the correct studies and the barrel and bullet variable labels to the new CSAFE labeling conventions. 
5. Add letters to the barrel, bullet, and land labels to conform to the CSAFE labeling conventions.
6. Create land ID variables that combine study, barrel, bullet, and land labels. 
7. Add a case variable (just numbering comparisons 1 through `r dim(bullet_train_raw)[1]` based on current order in the data). 
8. Reorder the variables.

@@

bullet_train <- 
  bullet_train_raw %>%
  # 1. Rename variables
  rename(
    "study1" = "study.x", "barrel1" = "barrel.x", "bullet1" = "bullet.x", "land1" = "land.x",
    "study2" = "study.y", "barrel2" = "barrel.y", "bullet2" = "bullet.y", "land2" = "land.y",
    "same_source"= "match"
  ) %>%
  # 2. Select variables of interest
  select(
    study1, barrel1, bullet1, land1,
    study2, barrel2, bullet2, land2,
    all_of(rtrees_features), same_source
  ) %>%
  # 3. Convert categorical variables to characters
  mutate(across(c(barrel1, bullet1, land1, barrel2, bullet2, land2), as.character)) %>%
  # 4. Adjust study, barrel, and bullet labels
  mutate(
    study1 = ifelse(study1 == "Hamby44", "Hamby173", study1),
    study2 = ifelse(study2 == "Hamby44", "Hamby173", study2),
    bullet1 = ifelse(barrel1 %in% letters, barrel1, bullet1),
    bullet2 = ifelse(barrel2 %in% letters, barrel2, bullet2),
    barrel1 = ifelse(barrel1 %in% letters, "Unk", barrel1),
    barrel2 = ifelse(barrel2 %in% letters, "Unk", barrel2)
    ) %>%
  # 5. Add abbreviations infront of barrel, bullet, and land labels
  mutate(
    barrel1 = paste0("Br", barrel1),
    barrel2 = paste0("Br", barrel2),
    bullet1 = paste0("B", bullet1),
    bullet2 = paste0("B", bullet2),
    land1 = paste0("L", land1),
    land2 = paste0("L", land2)
  ) %>%
  # 6. Create land ID variables
  mutate(
    land_id1 = paste(study1, barrel1, bullet1, land1, sep = "-"),
    land_id2 = paste(study2, barrel2, bullet2, land2, sep = "-")
  ) %>%
  # 7. Add a case variable
  mutate(case = as.character(1:n())) %>%
  # 8. Reorder the variables
  select(land_id1, land_id2, 
         study1, barrel1, bullet1, study2, land1, barrel2, bullet2, land2,
         all_of(rtrees_features), same_source)
```

@@

The updated signature identification labels are included below:

| Descriptor | Labels | 
| --- | -------------- |
| Study | `r unique(c(bullet_train$study1, bullet_train$study2))` |
| Barrel | `r unique(c(bullet_train$barrel1, bullet_train$barrel2))` |
| Bullet | `r unique(c(bullet_train$bullet1, bullet_train$bullet2))` |
| Land | `r unique(c(bullet_train$land1, bullet_train$land2))` |

>>>>

1. Rename the label variables using numbers instead of letters (e.g. `study.x` and `study.y` become `study1` and `study2`, respectively), and rename `match` as `samesource`. 
2. Select only the variables of interest for the study: signature ID labels (`study1`, `barrel1`, `bullet1`, `land1`, `study2`, `barrel2`, `bullet2`, and `land2`), features used to train the `rtrees` random forest model (`ccf`, `rough_cor`, `D`, `sd_D`, `matches`, `mismatches`, `cms`, `non_cms`, and `sum_peaks`), and the variable identifying whether the two signatures are fired from the same gun (`samesource`).
3. Convert all categorical variables to characters.
4. Change the study labels to the correct studies and the barrel and bullet variable labels to the new CSAFE labeling conventions. 
5. Add letters to the barrel, bullet, and land labels to conform to the CSAFE labeling conventions.
6. Create land ID variables that combine study, barrel, bullet, and land labels. 
7. Add a case variable (just numbering comparisons 1 through `r dim(bullet_train_raw)[1]` based on current order in the data). 
8. Select final variables for data to use in paper: number assinged to signature (`case`), land IDs (`land_ID1`, `land_ID2`), features used to train the `rtrees` random forest model (`ccf`, `rough_cor`, `D`, `sd_D`, `matches`, `mismatches`, `cms`, `non_cms`, and `sum_peaks`), and the variable identifying whether the two signatures are fired from the same gun (`samesource`).

@@

bullet_train <- 
  bullet_train_raw %>%
  # 1. Rename variables
  rename(
    "study1" = "study.x", "barrel1" = "barrel.x", "bullet1" = "bullet.x", "land1" = "land.x",
    "study2" = "study.y", "barrel2" = "barrel.y", "bullet2" = "bullet.y", "land2" = "land.y",
    "samesource"= "match"
  ) %>%
  # 2. Select variables of interest
  select(
    study1, barrel1, bullet1, land1,
    study2, barrel2, bullet2, land2,
    all_of(rtrees_features), samesource
  ) %>%
  # 3. Convert categorical variables to characters
  mutate(across(c(barrel1, bullet1, land1, barrel2, bullet2, land2), as.character)) %>%
  # 4. Adjust study, barrel, and bullet labels
  mutate(
    study1 = ifelse(study1 == "Hamby44", "Hamby173", study1),
    study2 = ifelse(study2 == "Hamby44", "Hamby173", study2),
    bullet1 = ifelse(barrel1 %in% letters, barrel1, bullet1),
    bullet2 = ifelse(barrel2 %in% letters, barrel2, bullet2),
    barrel1 = ifelse(barrel1 %in% letters, "Unk", barrel1),
    barrel2 = ifelse(barrel2 %in% letters, "Unk", barrel2)
    ) %>%
  # 5. Add abbreviations in front of barrel, bullet, and land labels
  mutate(
    barrel1 = paste0("Br", barrel1),
    barrel2 = paste0("Br", barrel2),
    bullet1 = paste0("B", bullet1),
    bullet2 = paste0("B", bullet2),
    land1 = paste0("L", land1),
    land2 = paste0("L", land2)
  ) %>%
  # 6. Create land ID variables
  mutate(
    land_id1 = paste(study1, barrel1, bullet1, land1, sep = "-"),
    land_id2 = paste(study2, barrel2, bullet2, land2, sep = "-")
  ) %>%
  # 7. Add a case variable
  mutate(case = as.character(1:n())) %>%
  # 8. Select final variables
  select(case, land_id1, land_id2, all_of(rtrees_features), samesource)
```

@@

The land IDs in the cleaned bullet training data are separated in the code below, so that the updated signature identification labels can be included in the table below:

```{r}
bullet_train_expanded <- 
  bullet_train %>% 
  separate(col = land_id1, into = c("study1", "barrel1", "bullet1", "land1")) %>%
  separate(col = land_id2, into = c("study2", "barrel2", "bullet2", "land2"))
```

| Descriptor | Labels | 
| --- | -------------- |
| Study | `r unique(c(bullet_train_expanded$study1, bullet_train_expanded$study2))` |
| Barrel | `r unique(c(bullet_train_expanded$barrel1, bullet_train_expanded$barrel2))` |
| Bullet | `r unique(c(bullet_train_expanded$bullet1, bullet_train_expanded$bullet2))` |
| Land | `r unique(c(bullet_train_expanded$land1, bullet_train_expanded$land2))` |
